Automated Formal Synthesis of Digital Controllers for State-Space Physical Plants

### 2017-01-25 cav2017/cav2017.tex
moved table "Benchmarks for State-Space Physical Plants" to appendix

\newcommand{\addtodo}[1]{{\color{red} TODO: #1}}
\newcommand{\commentib}[1]{{\color{blue} [IB: #1]}}
\newcommand{\mat}[1]{\boldsymbol{#1}}
\newcommand{\xmark}{\ding{55}}

@@

We describe and evaluate two approaches for synthesizing digital controllers for state-space physical plants, 
both based on the CounterExample Guided Inductive Synthesis (CEGIS)~\cite{sketch} architecture, recently of wide use for program synthesis.   
We provide two alternative automated procedures to synthesise safe digital state-feedback controls, 
and alongside completeness \todo{are we actually complete? for instance, abstraction refinement might not be successful (we do TO)} we prove their soundness by encompassing errors due to digitalisation and quantisation effects that are present in the digital controller interacting with the continuous plant. 
% computations and representations of dynamics, providing sufficient conditions for sound and complete safe solutions for the state feedback control problem.

@@

We provide experimental results 
showing that both approaches are able to efficiently synthesize stable and safe controllers 
for a set of intricate physical plant models taken from the digital 
control literature: the median runtime for our benchmark
set is $7.9$ seconds, and most controllers can be synthesized in less
than $17.2$ seconds.

@@

\todo{Have we clarified anywhere the time domain of validity of our claims? CT vs DT? }

@@

Furthermore, it is practically relevant to consider constraints on the input
signal $u_{k}$ and bounds on the initial states $x_0$,
which we assume have given bounds. %%  $\overline{u}$, $\underline{u}$
%% and $\overline{x_0}$, $\underline{x_0}$, respectively.
We refer to these constraints as $\phi_{init}$: \todo{input constraints should not be called 'init'}
$$\phi_{init} = {\forall k.\underline{u} \leq u_{k} \leq \overline{u}} 
\wedge \bigwedge_{i=1}^{n} \underline{x_{i,0}} \leq x_{i,0} \leq \overline{x_{i,0}}.$$  
In the first case, this means that the control input saturates in view of
physical constraints.

@@

Previous studies~\cite{gangli1} show that the FWL affects the poles and 
zeros positions, degrading the closed-loop dynamics, causing steady-state 
errors (see the Appendix~\ref{sec:appendix:LTIbackground} for details) 
and eventually destabilizing  the system~\cite{Bessa16}. However, since 
in this paper we require stability only as a precursor to safety, it is 
sufficient to check that the (perturbed, noisy) model converges to a neighbourhood of the equilibrium within the safe set.

@@

To argue for this sufficient check, 
consider the set of initial states, $X_0$, which we assume to be convex since it's a hypercube. 
Name $v_i$ its vertexes, where $i=1,\ldots, 2^n$.  
Thus any point $x \in X_0$ can be expressed by convexity as $x = \sum_{i=1}^{2^n} \alpha_i v_i$, 
where $\sum_{i=1}^{2^n} \alpha_i =1$. Then if $x_0=x$, we obtain 
\begin{align*}
x_k   &= (A_d - B_d K)^k x = (A_d - B_d K)^k \sum_{i=1}^{2^n} \alpha_i v_i 
      = \sum_{i=1}^{2^n} \alpha_i (A_d - B_d K)^k v_i = \sum_{i=1}^{2^n} \alpha_i x_k^i, 
 \end{align*}     
where $x_k^i$ denotes the trajectories obtained from the single vertex $v_i$. 
We conclude that any $k$-step trajectory is encompassed, within a convex set, by those generated from the vertices.

@@

\item The third {\sc verify} stage, {\sc complete}, checks that the
current $k$ is large enough to ensure that the boundaries are not
violated for any $k'{>}k$.  For this purpose, we compute the
completeness threshold $\overline{k}$ for the current candidate $K$
and check that $k{\leq}\overline{k}$. This is done according to the following argument, depicted in Fig \ref{fig:ct}.

A stable control system is known to have converging dynamics. 
Assume the closed-loop matrix eigenvalues are not repeated (which is sensible to do, since we select them). 
The distance of trajectories from the reference point (in this case the origin) decreases over time within subspaces related to real-valued eigenvalues;  
however, this in general not the case when dealing with complex eigenvalues.  
Consider the closed-loop matrix, which updates the states every discrete time step, 
and select the eigenvalue with smallest (non trivial) imaginary value -- call such value $\vartheta$. 
At every consecutive time steps $kT_s, (k+1)T_s$ ($T_s$ being the sampling time), 
the dynamics projected on the corresponding eigenspace rotate of $\vartheta T_s$ radiants.  
So taking $k$ as the ceiling of $\frac{2\pi}{\vartheta T_s}$, 
surely after $k$ steps we have completed a full rotation, which results in a point that is closer to the origin. 
This $k$ is then taken to be the completeness threshold.

\begin{figure*}[t]
\centering
\includegraphics[width=0.7\textwidth]{ct.eps}
\vspace{0.1cm}
\caption{Completeness threshold for multi-staged verification.}
\label{fig:ct}
\end{figure*}

@@

A set of state-space models for different classes of 
systems were extracted from the 
literature~\cite{Franklin15,maglev,converters,CTMS} 
and employed for validating our methodology. 
%All systems are SISO models of the following form:
%
%\begin{equation}
%\left\lbrace\begin{array}{c}
%\dot{\textbf{x}}(t)=\textbf{A}\textbf{x}(t)+\textbf{B}u(t)\\
%y(t)=\textbf{C}\textbf{x}(t)+\textbf{D}u(t)
%\end{array}\right.
%\end{equation}
%
Table~\ref{table:State-space benchmarks} summarizes 
all our benchmarks, which are SISO models (cf. section~\ref{sec:preliminaries}). In the Table~\ref{table:State-space benchmarks}, the Inverted Pendulum appears to be a two output system but it was treated as two SISO systems during the experiments. For the sake of simplicity all the states measurements are assumed available. In practice, state estimators should be designed in order to obtain an estimate of the states from the single output. Further extensions of this work can include state observers synthesis.

%\todo{they are not: we do not have a 1d output map (SO= single output), rather get access to all variables. so the output map (C) is the identity, if you will. 
%this is a good point to clarify this, mention possible extensions to observer-based feedback architectures. }

@@

All benchmarks were discretized with different sample times using 
the approach described in the Appendix~\ref{sec:appendix}. 
All experiments were performed considering $\underline{x_{i}}=-1$ and 
$\overline{x_{i}}=1$ and the inputs $u_{k}=0, \forall k>0$.
We have conducted the experimental evaluation on a 12-core 2.40 GHz
Intel Xeon E5-2440 with 96 GB of RAM and Linux OS. We used the Linux \emph{times} command
to measure CPU time used for each benchmark. The experimental runtime was limited to
one hour per benchmark.
%

@@

\label{table:State-space benchmarks}

@@

LTI models in the form: 
%
\begin{align}
\dot{\vec{x}}(t)&=\mat{A}\vec{x}(t)+\mat{B}\vec{u}(t)\ \ :\ \ \vec{x} \in \mathbb{R}^{n}, \vec{u} \in \mathbb{R}^m, \mat{A} \in \mathbb{R}^{n \times n},\mat{B} \in \mathbb{R}^{n \times m}\\
\vec{y}(t)&=\mat{C}\vec{x}(t)+\mat{D}\vec{u}(t)\ \ :\ \ \vec{y} \in \mathbb{R}^{o}, \mat{C} \in \mathbb{R}^{o \times n}, \mat{D}  \in \mathbb{R}^{o \times m}\nonumber
\end{align}
\noindent where $\dot{\vec{x}}(t)$ describes the state evolution equation; 
$\vec{y}(t)$ represents an instantaneous output equation; and $\mat{A}$, $\mat{B}$, $\mat{C}$, and $\mat{D}$ are matrices that fully specify 
a continuous plant. Eq.~\eqref{eq:ode} is soundly discretized 
(as shown in the appendix) into
%
\begin{align}
\vec{x}_{k+1}&=\mat{A}_d\vec{x}_k+\mat{B}_d\vec{u}_k\\
\vec{y}_k&=\mat{C}_d\vec{x}_k+\mat{D}_d\vec{u}_k .\nonumber
\end{align}
%
\end{document}

>>>>

\renewcommand{\baselinestretch}{0.975}
\newcommand{\addtodo}[1]{{\color{red} TODO: #1}}
\newcommand{\commentib}[1]{{\color{blue} [IB: #1]}}
\newcommand{\mat}[1]{\boldsymbol{#1}}
\newcommand{\xmark}{\ding{55}}

@@

We describe and evaluate two approaches for synthesizing digital controllers for state-space physical plants, 
both based on the CounterExample Guided Inductive Synthesis (CEGIS)~\cite{sketch} architecture, recently of wide use for program synthesis.   
We provide two alternative automated procedures to synthesise safe digital state-feedback controls, 
and alongside completeness we prove their soundness by encompassing errors due to digitalisation and quantisation effects that are present in the digital controller interacting with the continuous plant. 
% computations and representations of dynamics, providing sufficient conditions for sound and complete safe solutions for the state feedback control problem.

@@

We provide experimental results 
showing that both approaches are able to efficiently synthesize stable and safe controllers 
for a set of intricate physical plant models taken from the digital 
control literature: the median runtime for our benchmark
set is $7.9$s, i.e., most controllers can be synthesized in less
than $17.2$ seconds.

@@

@@

Furthermore, it is practically relevant to consider constraints on the input
signal $u_{k}$ and bounds on the initial states $x_0$,
which we assume have given bounds. %%  $\overline{u}$, $\underline{u}$
%% and $\overline{x_0}$, $\underline{x_0}$, respectively.
We refer to these constraints as $\phi_{init}$:
$$\phi_{init} = {\forall k.\underline{u} \leq u_{k} \leq \overline{u}} 
\wedge \bigwedge_{i=1}^{n} \underline{x_{i,0}} \leq x_{i,0} \leq \overline{x_{i,0}}.$$  
In the first case, this means that the control input saturates in view of
physical constraints.

@@

Previous studies~\cite{gangli1} show that the FWL affects the poles and 
zeros positions, degrading the closed-loop dynamics, causing steady-state 
errors (see the Appendix~\ref{sec:appendix:LTIbackground} for details) 
and eventually destabilizing  the system~\cite{Bessa16}. However, since 
in this paper we require stability only as a precursor to safety, it is 
sufficient to check the (perturbed, noisy) model converges to a point 
within the safe states.

@@

To argue for this sufficient check, 
consider the set of initial states, $X_0$, which we assume to be convex since it's a hypercube. 
Name $v_i$ its vertexes, where $i=1,\ldots, 2^n$.  
Thus any point $x \in X_0$ can be expressed by convexity as $x = \sum_{i=1}^{2^n} \alpha_i v_i$, 
where $\sum_{i=1}^{2^n} \alpha_i =1$. Then if $x_0=x$, we obtain 
\begin{align*}
x_k   &= (A_d - B_d K)^k x = (A_d - B_d K)^k \sum_{i=1}^{2^n} \alpha_i v_i \\
      &= \sum_{i=1}^{2^n} \alpha_i (A_d - B_d K)^k v_i = \sum_{i=1}^{2^n} \alpha_i x_k^i, 
 \end{align*}     
where $x_k^i$ denotes the trajectories obtained from the single vertexes $v_i$.

@@

\item The third {\sc verify} stage, {\sc complete}, checks that the
current $k$ is large enough to ensure that the boundaries are not
violated for any $k'{>}k$.  For this purpose, we compute the
completeness threshold $\overline{k}$ for the current candidate $K$
and check that $k{\leq}\overline{k}$. This is done according to the following argument.

A stable control system is known to have converging dynamics. 
Assume the closed-loop matrix eigenvalues are not repeated (which is sensible to do, since we select them). 
The distance of trajectories from the reference point (in this case the origin) decreases over time within subspaces related to real-valued eigenvalues. 
However, this in general not the case when dealing with complex eigenvalues.  
Consider the matrix $(A_d - B_d K)$, which updates the states every time step, and select eigenvalue with smallest (non trivial) imaginary value.  
call such value $\vartheta$. 
At every time interval $kT_s, (k+1)T_s$ ($T_s$ being the sampling time), 
the dynamics projected on the corresponding eigenspace rotate of $\vartheta T_s$ radiants.  
So surely taking $k$ as the ceiling of $\frac{2\pi}{\vartheta T_s}$, 
after $k$ steps we have completed a rotation. 
This is the completeness threshold.

\begin{figure*}[t]
\centering
\includegraphics[width=0.7\textwidth]{ct.eps}
\vspace{0.1cm}
\caption{Completeness threshold for multi-staged verification.}
\end{figure*}

@@

A set of state-space models for different classes of 
systems were extracted from the 
literature~\cite{Franklin15,maglev,converters,CTMS} 
and employed for validating our methodology.

@@

Section~\ref{sec:BenchmarksforStateSpace} in the appendix summarizes 
all our benchmarks, which are SISO models (cf. section~\ref{sec:preliminaries}). 
The Inverted Pendulum appears to be a two output system, but it was treated as two 
SISO systems during the experiments. For the sake of simplicity all the states 
measurements are assumed available. In practice, state estimators should be designed 
in order to obtain an estimate of the states from the single output. 
Further extensions of this work can include 
state observers synthesis.

All benchmarks were discretized with different sample times using 
the approach described in the Appendix~\ref{sec:appendix:LTIbackground}. 
All experiments were performed considering $\underline{x_{i}}=-1$ and 
$\overline{x_{i}}=1$ and the inputs $u_{k}=0, \forall k>0$.
We have conducted the experimental evaluation on a 12-core 2.40 GHz
Intel Xeon E5-2440 with 96 GB of RAM and Linux OS. We used the Linux \emph{times} command
to measure CPU time used for each benchmark. The experimental runtime was limited to
one hour per benchmark.
%

LTI models in the form: 
%
\begin{align}
\dot{\vec{x}}(t)&=\mat{A}\vec{x}(t)+\mat{B}\vec{u}(t)\ \ :\ \ \vec{x} \in \mathbb{R}^{n}, \vec{u} \in \mathbb{R}^m, \mat{A} \in \mathbb{R}^{n \times n},\mat{B} \in \mathbb{R}^{n \times m}\\
\vec{y}(t)&=\mat{C}\vec{x}(t)+\mat{D}\vec{u}(t)\ \ :\ \ \vec{y} \in \mathbb{R}^{o}, \mat{C} \in \mathbb{R}^{o \times n}, \mat{D}  \in \mathbb{R}^{o \times m}\nonumber
\end{align}
\noindent where $\dot{\vec{x}}(t)$ describes the state evolution equation; 
$\vec{y}(t)$ represents an instantaneous output equation; and $\mat{A}$, $\mat{B}$, $\mat{C}$, and $\mat{D}$ are matrices that fully specify 
a continuous plant. Eq.~\eqref{eq:ode} is soundly discretized 
(as shown in the appendix) into
%
\begin{align}
\vec{x}_{k+1}&=\mat{A}_d\vec{x}_k+\mat{B}_d\vec{u}_k\\
\vec{y}_k&=\mat{C}_d\vec{x}_k+\mat{D}_d\vec{u}_k .\nonumber
\end{align}
%

%-------------------------------
\section{Benchmarks for State-Space Physical Plants}
\label{sec:BenchmarksforStateSpace}
%-------------------------------

@@

\label{table:State-space-benchmarks}

@@

\end{document}
